---
title: ""
date: ""
output: 
  html_document:
    toc: true
    toc_float: true
    css: "style.css"
    includes:
      in_header: header_igua.html
---

## Zenodo repository

All files (.tsv/.RDS) referenced in this notebook can be downloaded from https://zenodo.org/uploads/15394038. Please note that for the supplementary tables, it is assumed that the headers have been removed prior to importing the file in R.

## Supplementary tables

* Supplementary table 1: The manually curated dataset of BGCs, originally from the BiG-SCAPE paper
* Supplementary table 2: BiG-SCAPE clustering of the manually curated dataset using glocal clustering
* Supplementary table 3: BiG-SCAPE clustering of the manually curated dataset using global clustering
* Supplementary table 4: IGUA clustering of the manually curated dataset with default settings
* Supplementary table 5: BiG-SCAPE clustering of the *Streptomyces* dataset using glocal clustering
* Supplementary table 6: IGUA clustering of the *Streptomyces* dataset with default settings
* Supplementary table 7: *Streptomyces* samples described
* Supplementary table 8: Time taken to cluster *Streptomyces* samples by IGUA
* Supplementary table 9: proG3 samples described
* Supplementary table 10: Speed and time used to cluster proG3 samples by IGUA, BiG-SCAPE and BiG-SLiCE
* Supplementary table 11: IGUA clustering of the proG3 dataset with default settings
* Supplementary table 12: Metadata attached to the proG3 Seurat object
* Supplementary table 13: IGUA clustering of secretion system dataset with default settings
* Supplementary table 14: GTDB assignments for secretion system host genomes
* Supplementary table 15: Metadata attached to the secretion systems Seurat object
* Supplementary table 16: IGUA clustering of the prophage dataset with default settings
* Supplementary table 17: IGUA clustering of the prophage dataset with 0.99 threshold
* Supplementary table 18: Metadata attached to the prophage Seurat object

## Other tables used in this notebook

* `prophage_metadata.tsv` - metadata to attach to the prophage Seurat object, leading to supplementary table 18
* `single_locus_clusters_renamed2tax.tsv` - taxonomic assignments of the secretion system gene clusters, similar to supplementary table 14
* `single_locus_clusters_renamed.tsv` - additional metadata for the secretion systems, included in the larger supplementary table 15
* `bgcs_to_phylum.tsv` - maps BGCs in the proG3 dataset to taxonomic information
* `AMI_benchmark_data.tsv`- AMI values for each clustering distance and linkage method tested when optimising default settings for IGUA
* `NMI_benchmark_data.tsv` - NMI values for each clustering distance and linkage method tested when optimising default settings for IGUA
* `purity_benchmark_data.tsv` - purity values for each clustering distance and linkage method tested when optimising default settings for IGUA
* `GCFs_benchmark_data.tsv` - the number of GCFs created for each clustering distance and linkage method tested when optimising default settings for IGUA
* `41589_2019_400_MOESM31_ESM.xlsx` - original Excel file from the BiG-SCAPE paper detailing their manually curated dataset. This can be downloaded from https://static-content.springer.com/esm/art%3A10.1038%2Fs41589-019-0400-9/MediaObjects/41589_2019_400_MOESM31_ESM.xlsx.
* `proG3_labels.tsv` - labels associated with the Pfam annotation of the proG3 dataset
* `biosynthetic_classes.tsv` - the biosynthetic class annotation of the BGCs from the proG3 dataset
* `prophage_labels.tsv` - labels associated with the Pfam annotation of the prophage dataset
* `sec_labels.tsv` - labels associated with the Pfam annotation of the secretion system dataset

## Seurat objects

* `pfams_proG3.RDS` - Seurat object associated with the proG3 BGC dataset
* `pfams_secretion.RDS`- Seurat object associated with the secretion systems dataset
* `pfams_prophage.RDS` - Seurat object associated with the prophage dataset

## Additional scripts and code

* `proG3_rep_sampling.R` was used to create the proGenomes3 samples used for speed and memory testing. This script uses the file `proGenomes3_specI_lineageGTDB.tab`, downloaded from https://progenomes.embl.de/download.cgi. 
* `ìgua_benchmark_parameter.R` was used to generate the AMI, NRI, NMI, purity and GCF counts for the benchmarking done using the manually curated dataset. This script was re-used for the purity measurements calculated, and thus some lines of code are commented out depending on what was calculated. In its current state, the script creates the GCF dataframe. 
* `strepto_sampling.py` was used to create the *Streptomyces* samples.
* `umap_bgcs.py` was used to annotate Pfam domains among GCF representatives for the proG3 BGC dataset.
* `umap_prophage.py` was used to annotate Pfam domains among GCF representatives for the prophage dataset.
* `umap_secretion.py` was used to annotate Pfam domains among GCF representatives for the secretion systems dataset.

The command used to run BiG-SCAPE (v1.1.5) on the benchmark dataset is shown below (only showing global mode). The `$variable` represented a list of the manually curated dataset BGC IDs, to only include those BGCs in the clustering.

```{r, results='hide', eval=FALSE, echo=TRUE}
bigscape.py --label benchmark --inputdir /MIBiG_1.3/MIBiG_1.3_final --include_gbk_str $variable --outputdir /big_scape_benchmark/default_global/ --cores 8 --mode global --pfam_dir /databases/pfam --anchorfile /anchor_domains.txt --cutoffs 0.75 --hybrids-off
```

The command used to run BiG-SCAPE (v1.1.5) on the *Streptomyces* dataset is shown below.

```{r, results='hide', eval=FALSE, echo=TRUE}
bigscape.py --label strepto --inputdir /strepto_results/ --outputdir /big_results/ --cores 28 --mode glocal --pfam_dir /databases/pfam --anchorfile /anchor_domains.txt --include_singletons --mibig --hybrids-off
```

An example command used to run BiG-SLiCE v(1.1.0) on the proGenomes3 representative samples is shown below. This was used in a Nextflow script (modified to work with Nextflow) to create report logging time and memory usage.

```{r, results='hide', eval=FALSE, echo=TRUE}
bigslice -i input_sample_4 --threshold 300 -t 1 output_input_sample_4
```

An example command used to run BiG-SCAPE (v1.1.5) on the proGenomes3 representative samples is shown below. This is was used in a Nextflow script (modified to work with Nextflow) to create reports logging time and memory usage.

```{r, results='hide', eval=FALSE, echo=TRUE}
bigscape.py --label sample_1_8_cpus --inputdir /parsed_results --include_gbk_str $variable --outputdir sample_1_8_cpus --cores 8 --mode glocal --pfam_dir /databases/pfam --anchorfile /anchor_domains.txt --include_singletons --mibig --hybrids-off
```


## Figure code

Code for the figures has been included below. Please note that the chunks are not executed and thus figures will not be included in the knitted notebook, only the code chunks. Also note that any time a supplementary table is referred to, it is assumed that the table header has been removed prior.

```{r, results='hide', eval=FALSE, echo=TRUE}

#Packages used throughout this notebook

library(readxl)
library(tidyverse)
library(khroma)
library(ggnewscale)
library(ggpubr)
library(ComplexHeatmap)
library(ggplotify)
library(ggalluvial)
library(ggmagnify)
library(gridExtra)
library(ggsignif)
library(Signac)
library(Seurat)
library(clustree)

```

```{r, eval=FALSE, echo=TRUE, results='hide'}
###FIGURE 2 MASTER CODE

##Panel B

#Importing and fixing the benchmark dataset dataframe

curated=read_excel("41589_2019_400_MOESM31_ESM.xlsx")
curated=curated[-c(288,289,290,291,292,293,294,308,309,310),]
sep_curated=curated %>%
  separate_rows(`MIBiG accession`, sep = ",")
sep_curated$`MIBiG accession`=str_trim(sep_curated$`MIBiG accession`, side=c("both"))
sep_curated=data.frame(sep_curated$`MIBiG accession`,sep_curated$Group)
colnames(sep_curated)=c("mibig","group")
ann_curated=data.frame(curated$Class, curated$Group)
colnames(ann_curated)=c("class","group")
BGCs=c("BGC0000015", "BGC0000095","BGC0000322",
       "BGC0000462","BGC0000629","BGC0000875",
       "BGC0000945", "BGC0001081", "BGC0001147",
       "BGC0001208", "BGC0001232", "BGC0001347")
b=length(BGCs)
temp1=0
for (y in 1:b){
  temp1=c(temp1,c(which(sep_curated$mibig==BGCs[y])))
}
sep_curated=sep_curated[-temp1,]

#Importing IGUA clustering at default settings

benchmark=read.delim(paste0("supplementary_table_S4.tsv"))
benchmark$cluster_id=strtrim(benchmark$cluster_id,10)
benchmark=benchmark %>%
  group_by(gcf_id) %>%
  filter(n() >=2)
igua=data.frame(benchmark$cluster_id,benchmark$gcf_id)
colnames(igua)=c("mibig","GCF")
BGCs=c("BGC0000015", "BGC0000095","BGC0000322",
       "BGC0000462","BGC0000629","BGC0000875",
       "BGC0000945", "BGC0001081", "BGC0001147",
       "BGC0001208", "BGC0001232", "BGC0001347")
b=length(BGCs)
temp=0
for (y in 1:b){
  temp=c(temp,c(which(igua$mibig==BGCs[y])))
}
igua_f=igua[-temp,]

#Merging the two dataframes

merged_data=merge(sep_curated,igua_f,by="mibig", all.x = TRUE)

data2=merged_data %>%
  group_by(group, GCF) %>% 
  summarise(n=n()) 

data2$row=row.names(data2)

#Creating the biosynthetic class annotation

ann_merge=data2 %>% inner_join(ann_curated,
                               by="group")

ann_merge=ann_merge[!duplicated(ann_merge$row),]
ann_row=data.frame(ann_merge$group,ann_merge$class)
ann_row=ann_row[!duplicated(ann_row$ann_merge.group),]
row.names(ann_row)=ann_row$ann_merge.group
ann_row2=data.frame(ann_row$ann_merge.class)
row.names(ann_row2)=row.names(ann_row)
colnames(ann_row2)=c("Class")

ann_colors = list(
  Class = c(Alkaloids = "#EF6C00", NRPs = "#884EA0", `Other Hybrids` = "#607D8B",
            Others = "grey", Polyketides = "#1E88E5", `Polyketides/NRPs` = "palegreen",
            RiPPs = "#FDD835", Saccharides = "#EC407A", Terpenes = "#009688"))

#Sorting the heatmap contingency table using a custom function

contingency.table <- table(merged_data$GCF, merged_data$group)
my.truth <- names(sort(table(merged_data$group), decreasing = T))
gecco.clusters <- names(sort(table(merged_data$GCF), decreasing = T))
yeet <- function(mymat, ref){
  find.clusters <- c()
  used.clusters <-c()
  # loop through reference list
  for (i in 1:length(ref)){
    clus <- ref[i]
    # take the column of the contintingency table that matches the current reference GCF
    # finds the row that yields the GCF with maximum overlap
    # if there are ties, R takes the first one
    clus.bgcs <- rownames(mymat)[which.max(mymat[,which(colnames(mymat)==clus)])]
    # if that particular maximum GCF hasn't been included yet...
    # (this is done to deal with ties)
    if (!(clus.bgcs%in%used.clusters)){
      # append the maximum GCF to our final list of clusters, as well as out "used" GCF list
      find.clusters <- c(find.clusters, clus.bgcs)
      used.clusters <- c(used.clusters, clus.bgcs)
    }
  }
  for (i in 1:length(rownames(mymat))){
    clus.missing <- rownames(mymat)[i]
    # if the GCF isn't included in our final list...
    if (!(clus.missing%in%used.clusters)){
      # add it
      used.clusters <- c(used.clusters, clus.missing)
      find.clusters <- c(find.clusters, clus.missing)
    }
  }
  # return ordered GCFs
  return(find.clusters)
}
order.big <- yeet(mymat = contingency.table, ref = my.truth)
final <- contingency.table
final <- final[match(order.big, rownames(final)),]
order.gecco <- yeet(mymat = t(final), ref = rownames(final))
final <- final[,match(order.gecco, colnames(final))]

#Setting the colour palette with 0 set to a charcoal colour

bgc_palette <- hcl.colors(50, "YlGnBu")
bgc_palette[1] <- "#171717"

#Creating the annotated heatmap

p6h <- pheatmap(t(final), color = bgc_palette, border_color = F,cluster_rows = F,
                      cluster_cols = F,
                      show_colnames = FALSE,fontsize = 6, annotation_row = ann_row2,
                      annotation_colors = ann_colors, annotation_names_row = FALSE,
                      heatmap_legend_param = list(title = "# of BGCs", at = c(0,5,10,15),
                                                  direction = "vertical"))

#Converting the heatmap to a ggplot object

p6 = as.ggplot(p6h)

##Panel A

#Setting the colour palette and importing the ARI data

bright <- color("bright")

ARI_benchmark_data <- read.csv("ARI_benchmark_data.tsv", sep="")

#Making the ARI line plot

p5=ggplot(ARI_benchmark_data,aes(x=distance,y=ARIs,group=group, color=group))+
  geom_line(lwd=1)+ylab("ARI")+xlab("Clustering distance")+
  theme_classic()+
  scale_color_manual(values=c(bright(7)))+  
  labs(color="IGUA Linkage method")+
  # guides(color=guide_legend(ncol=2)) +
  new_scale_color()+
  geom_hline(aes(yintercept=0.7725007,
                 color="glocal"), linetype=2)+
  geom_hline(aes(yintercept=0.7520675,
                 color="global"), linetype=3)+
  scale_color_manual(values=c("black", "black"))+
  labs(color="BiG-SCAPE")+
  theme(axis.title = element_text(size=15),
        axis.text = element_text(size=13.5),
        legend.text = element_text(size=13.5),
        legend.title = element_text(size=15))+
  annotate("Text", x=0.8,
           y=0.90011233, label="X", size=6, color="red")

##Panels C and D

#Creating the dataframe

sample=c(rep(c(1:5),9))
method=c(rep("BiG-SCAPE",5),rep("IGUA",5), rep("BiG-SLiCE",5),
         rep("BiG-SCAPE",5), rep("IGUA",5), rep("BiG-SLiCE",5), rep("BiG-SCAPE",5),
         rep("IGUA",5), rep("BiG-SLiCE",5))
BGCs=c(rep(c(11478, 11908, 11214, 
             11227, 11144),6))
cpu=c(rep(1,15), rep(8,15), rep(28,15))
memory=c(18.035, 18.081, 19.118, 20.700, 19.834,
         2.046, 2.120, 2.025, 2.014, 2.004,
         3.037, 2.933, 3.009, 2.977, 3.077,
         58.227, 58.638, 55.205, 56.161, 52.465,
         2.243, 2.313, 2.221, 2.211, 2.198,
         8.779, 8.278, 8.893, 8.860, 8.908,
         168.613, 114.552, 160.314, 162.163, 152.919,
         2.727, 2.801, 2.703, 2.685, 2.682,
         14.811, 14.184, 14.980, 15.041, 14.695)
time=c(1560, 1740, 1560, 1560, 1560,
       34.85, 38.32, 30.8, 32.25, 32.97,
       331, 346, 320, 313, 317,
       362, 331, 353, 312, 316,
       4.83, 5.18, 4.75, 4.52, 4.58,
       54.233, 56.267, 52.45, 51.05, 51.917,
       244, 267, 265, 235, 212, 
       1.98, 2.12, 1.98, 1.9, 2.08,
       24.233, 25.783, 23.817, 23.45, 23.717)
df=data.frame(sample,method,cpu,time, memory)
df$cpu=as.factor(df$cpu)

#Making the barplots

p13=ggplot(df, aes(x=cpu, y=time, fill=method))+
  geom_bar(stat = "summary",
           position = "dodge",
           fun=mean)+
  geom_point(aes(y=time, group=method),
             stat="identity", position=position_dodge(width=0.9),
             pch=1, size=2, show.legend = FALSE)+
  theme_classic()+
  theme(legend.title = element_blank(),
        axis.title = element_text(size=15),
        legend.text = element_text(size=13.5),
        legend.position="inside",
        legend.justification.inside = c(0.98,0.98),
        axis.text = element_text(size=13.5))+
  ylab("Real time (min)")+
  xlab("CPUs")+
  scale_fill_manual(values=c("#33BBEE","#EE3377","#009988"))+
  geom_signif(
    y_position = c(1900, 1800), xmin = c(0.70, 1), xmax = c(1.25, 1.25),
    annotation = c("x47.2", "x9.6"), tip_length = 0,
    textsize = 3,
    size = 0.7
  )+
  geom_signif(
    y_position = c(600, 500), xmin = c(1.71, 2), xmax = c(2.285, 2.285),
    annotation = c("x70.2", "x11.1"), tip_length = 0,
    textsize = 3,
    size = 0.7
  )+
  geom_signif(
    y_position = c(550, 450), xmin = c(2.71, 3.02), xmax = c(3.29, 3.29),
    annotation = c("x121.6", "x12.0"), tip_length = 0,
    textsize = 3,
    size = 0.7
  )

p14=ggplot(df, aes(x=cpu, y=memory, fill=method))+
  geom_bar(stat = "summary",
           position = "dodge",
           fun=mean)+
  geom_point(aes(y=memory, group=method),
             stat="identity", position=position_dodge(width=0.9),
             pch=1, size=2, show.legend = FALSE)+
  theme_classic()+
  theme(legend.title = element_blank(),
        axis.title = element_text(size=15),
        legend.text = element_text(size=13.5),
        legend.position="inside",
        legend.justification.inside = c(0.08,0.98),
        axis.text = element_text(size=13.5))+
  ylab("Peak RSS (Gb)")+
  xlab("CPUs")+
  scale_fill_manual(values=c("#33BBEE","#EE3377","#009988"))+
  geom_signif(
    y_position = c(50, 40), xmin = c(0.70, 1), xmax = c(1.25, 1.25),
    annotation = c("x9.4", "x1.5"), tip_length = 0,
    textsize = 3,
    size = 0.7
  )+
  geom_signif(
    y_position = c(80, 70), xmin = c(1.71, 2), xmax = c(2.285, 2.285),
    annotation = c("x25.1", "x3.9"), tip_length = 0,
    textsize = 3,
    size = 0.7
  )+
  geom_signif(
    y_position = c(190, 180), xmin = c(2.71, 3.02), xmax = c(3.29, 3.29),
    annotation = c("x55.8", "x5.4"), tip_length = 0,
    textsize = 3,
    size = 0.7
  )

#Adjusting some figure elements to make a better panel figure

p5 <- p5 +
  theme(legend.position = "inside",
        legend.position.inside = c(0.25, 0.4),
        legend.background = element_rect(fill = NA),
        legend.text = element_text(size = 11.5),
        legend.title = element_text(size = 11.5, face = "bold"))

p13 <- p13 + theme(legend.position="none")
p14 <- p14 + theme(legend.background = element_rect(fill = NA),
                   legend.text = element_text(size = 11.5))
                   
#Arranging the panels

figure2_arrange <- ggarrange(
  ggarrange(p5,p13,p14, align = "hv", nrow = 3, ncol = 1, labels = c("A", "C", "D"), heights = c(3,2,2)),
  p6, align = "hv", ncol = 2, widths = c(2,5), labels = c("A","B")
)

#Saving the plot

ggsave("figure2.svg", plot = figure2_arrange, device = "svg", dpi = 300, height = 14.4, width = 16.8)
```



```{r, results='hide', eval=FALSE, echo=TRUE}
###FIGURE 3 MASTER CODE

#Setting the colour palette

precols = c("#15889C", "#ED496F", "#8E1382", "#FFB300", "#FF6F00", "#43A047",
            "#2D62A3", "#1F7530", "#573794", "#28AAE1", "#870F0F", "#24b177",
            "#DDDDDD", "#7DCEA0", "#85C1E9", "#EB984E", "#EAD9D5", "#03CD4A",
            "#CDDC39", "#E0594B", "#C76CDE", "#8D6E63", "#486FF7", "#6300B5",
            "#88E200", "#012824", "#0D3290", "#A347FB", "#54FC7A", "#EB1388",
            "#B0978D", "#FE52CF", "#83F1F6", "#F1F847", "#2B1DFC", "#6C6F15", 
            "#6CA05C", "#7788CD", "#F502F3", "#0DC290", "#FA0E03", "#3CAA0A",
            "#BEFC8D", "#08F8EB", "#B1CD3F", "#D6A5FA", "#CE606C", "#AB1EBA",
            "#6ECC9F", "#054DDC", "#854F49", "#F22B21", "#3A0E43", "#225805",
            "#37D160", "#E4B974", "#A8BADE", "#47EDD1", "#F47A92", "#9106EB",
            "#81AA20", "#D7FDFD", "#5DEB2E", "#F82745", "#6435E0", "#027FFE",
            "#8E3101", "#16F648", "#1C15BC", "#8BE46E", "#8D6FA0", "#E68FC6",
            "#058CA9", "#9E018A", "#BDFD0B", "#B22760", "#2BF49F", "#CB9348",
            "#9D8303", "#C251A1", "#46ADAF", "#A3E3AF", "#22BB34", "#6EA3FA",
            "#260374", "#1C3854", "#405D37", "#C21DF3", "#FCEA92", "#537F88",
            "#FD4C18", "#F2D71E")

##Prophage section

#importing the RDS file and creating the prophage UMAP figure
pfams_pro <- readRDS("pfams_prophage.RDS")
pro_umap=DimPlot(pfams_pro, cols=precols, label=FALSE, raster=FALSE, pt.size=1,
                 alpha = 0.5)+NoLegend()

#importing prophage clustering by IGUA and associated metadata
prophage_metadata <- read.delim("prophage_metadata.tsv", header=TRUE)
pro_igua=read.delim('supplementary_table_S17.tsv', header=TRUE)

colnames(prophage_metadata)[1]="cluster_id"

#merging to one master file
pro_metadata_igua=merge(prophage_metadata, pro_igua, by="cluster_id")

#filtering singleton GCFs
pro_metadata_igua= pro_metadata_igua %>% 
  group_by(gcf_id) %>%
  filter(n() >=2)

#making the histogram for phyla per GCFs
metadata_hist= pro_metadata_igua %>% 
  group_by(phylum, gcf_id) %>% 
  summarise(n=n())


histogram_data <- metadata_hist %>%
  group_by(gcf_id) %>%
  summarise(phylum_count = n_distinct(phylum))


pro_hist_phyla=ggplot(histogram_data, aes(x=phylum_count))+
  geom_histogram(fill="#ED496F", binwidth = 1)+
  xlab("Phyla per GCF")+
  ylab("# of GCFs")+
  theme_classic()+
  theme(axis.text = element_text(size=13.5),
        axis.title = element_text(size=15))+
  xlim(c(0,15))+
  scale_y_continuous(breaks = seq(0,15000,2500))

#making the second histogram, with gene clusters per GCF

pro_gcfs=pro_igua %>% 
  group_by(gcf_id) %>% 
  summarise(n=n())

pro_hist_GCs=ggplot(pro_gcfs, aes(x=n))+
  geom_histogram(fill="#15889C", binwidth = 1)+
  xlab("Gene clusters per GCF")+
  ylab("# of GCFs")+
  theme_classic()+
  theme(axis.text = element_text(size=13.5),
        axis.title = element_text(size=15))+
  xlim(c(0,15))+
  scale_y_continuous(breaks = seq(0,75000,25000))

#making the size reduction bar plot

Step=c(1:4)
n=c(100, (340994/356776)*100, (326220/356776)*100, (106749/356776)*100)

df_pro=data.frame(Step, n)

df_interp <- data.frame(
  Step = seq(1, 4, by = 0.0001)  # finer steps
) %>%
  mutate(n = approx(df_pro$Step, df_pro$n, xout = Step)$y)

df_interp$StepGroup <- cut(
  df_interp$Step,
  breaks = c(0, 1, 2, 3, 4),
  labels = c("GCs", "FRs", "NRs", "GCFRs"),
  right = TRUE
)

# Define colors for each StepGroup
MyColours <- precols[1:3]

# Plot
pro_area_size=ggplot() +
  geom_area(data = df_interp, aes(x = Step, y = n, fill = StepGroup),alpha=0.45,
            show.legend = FALSE) +
  geom_line(data = df_pro, aes(x = Step, y = n), colour = "grey60", size = 2) +
  geom_point(data = df_pro, aes(x = Step, y = n), colour = "grey60", size = 3) +
  geom_point(data = df_pro, aes(x = Step, y = n), colour = "white", size = 1.5) +
  scale_fill_manual(values = MyColours, name = "Step Intervals") +
  scale_x_continuous(breaks = 1:4, labels = c("GCs", "FRs", "NRs", "GCFRs")) +
  ylab("% of initial dataset size") +
  theme_classic() +
  theme(
    axis.title.x = element_blank(),
    axis.text = element_text(size = 13.5),
    axis.title = element_text(size = 15),
    legend.text = element_text(size = 13.5),
    legend.title = element_text(size = 15)
  )

##Secretion system section

#importing the RDS file and creating the secretion system UMAP

pfams_sec <- readRDS("pfams_secretion.RDS")

sec_umap=DimPlot(object = pfams_sec, label = FALSE, cols=precols, alpha=0.5,
                              pt.size = 1) + NoLegend()

#importing secretion system clustering by IGUA and associated metadata

sec_tax=read.delim("single_locus_clusters_renamed2tax.tsv", header=FALSE)

secretion_metadata <- read.delim("single_locus_clusters_renamed.tsv", header=FALSE)
sec_igua=read.delim('supplementary_table_S13.tsv', header=TRUE)

#setting up one master file for secretion systems

colnames(sec_igua)[1]="secretion"
colnames(secretion_metadata)=c("contig_id","secretion","start","end","major_secretion_system_type",
                               "minor_secretion_system_type","NCBI_nucleotide_accession")

metadata_igua=merge(secretion_metadata, sec_igua, by="secretion")

df_split <- do.call(rbind, strsplit(as.character(sec_tax$V4), ";"))
colnames(df_split) <- c("kingdom","phylum", "class","order", "family",
                        "genus","species")

df <- cbind(sec_tax, df_split)

colnames(df)[1:4]=c("secretion","NCBI_nucleotide_accession","gcf_XX","GTDBtax")

tax_metadata_igua=merge(df, metadata_igua, by="secretion")

#filtering singleton GCFs

tax_metadata_igua= tax_metadata_igua %>% 
  group_by(gcf_id) %>%
  filter(n() >=2)

#making the histogram for phyla per GCF

metadata_hist= tax_metadata_igua %>% 
  group_by(phylum, gcf_id) %>% 
  summarise(n=n())


histogram_data <- metadata_hist %>%
  group_by(gcf_id) %>%
  summarise(phylum_count = n_distinct(phylum))


sec_hist_phyla=ggplot(histogram_data, aes(x=phylum_count))+
  geom_histogram(fill="#ED496F", binwidth = 1)+
  xlab("Phyla per GCF")+
  ylab("# of GCFs")+
  theme_classic()+
  theme(axis.text = element_text(size=13.5),
        axis.title = element_text(size=15))+
  xlim(c(0,15))+
  scale_y_continuous(breaks = seq(0,900,200))

#making a second histogram for gene clusters per GCF

sec_gcfs=sec_igua %>% 
  group_by(gcf_id) %>% 
  summarise(n=n())


sec_hist_GCs=ggplot(sec_gcfs, aes(x=n))+
  geom_histogram(fill="#15889C", binwidth = 1)+
  xlab("Gene clusters per GCF")+
  ylab("# of GCFs")+
  theme_classic()+
  theme(axis.text = element_text(size=13.5),
        axis.title = element_text(size=15))+
  xlim(c(0,15))+
  scale_y_continuous(breaks = seq(0,2000,250))

#making the size reduction bar plot


Step=c(1:4)
n=c(100, (7321/10576)*100, (5512/10576)*100, (2744/10576)*100)

df_sec=data.frame(Step, n)

df_interp <- data.frame(
  Step = seq(1, 4, by = 0.0001)  # finer steps
) %>%
  mutate(n = approx(df_sec$Step, df_sec$n, xout = Step)$y)

df_interp$StepGroup <- cut(
  df_interp$Step,
  breaks = c(0, 1, 2, 3, 4),
  labels = c("GCs", "FRs", "NRs", "GCFRs"),
  right = TRUE
)

# Define colors for each StepGroup
MyColours <- precols[1:3]

# Plot
sec_area_size=ggplot() +
  geom_area(data = df_interp, aes(x = Step, y = n, fill = StepGroup),alpha=0.45,
            show.legend = FALSE) +
  geom_line(data = df_sec, aes(x = Step, y = n), colour = "grey60", size = 2) +
  geom_point(data = df_sec, aes(x = Step, y = n), colour = "grey60", size = 3) +
  geom_point(data = df_sec, aes(x = Step, y = n), colour = "white", size = 1.5) +
  scale_fill_manual(values = MyColours, name = "Step Intervals") +
  scale_x_continuous(breaks = 1:4, labels = c("GCs", "FRs", "NRs", "GCFRs")) +
  ylab("% of initial dataset size") +
  theme_classic() +
  theme(
    axis.title.x = element_blank(),
    axis.text = element_text(size = 13.5),
    axis.title = element_text(size = 15),
    legend.text = element_text(size = 13.5),
    legend.title = element_text(size = 15)
  )

##BGCs section (proG3 dataset)

#importing the RDS file and creating the secretion system UMAP

pfams <- readRDS("pfams_proG3.RDS")

bgcs_umap=DimPlot(object = pfams, label = FALSE, cols=precols, alpha=0.5, pt.size=1, 
            raster=FALSE) + NoLegend()

#importing BGC clustering by IGUA and associated metadata

proG3_igua=read.delim('supplementary_table_S11.tsv', header=TRUE)
bgcs_to_phylum=read.delim('bgcs_to_phylum.tsv', header=TRUE)

#filtering singleton GCFs from file with taxonomic information

bgcs_to_phylum = bgcs_to_phylum%>% 
  group_by(gcf_id) %>%
  filter(n() >=2)

#making histogram for phyla per GCF

metadata_hist_bgcs= bgcs_to_phylum%>% 
  group_by(phylum, gcf_id) %>% 
  summarise(n=n())

histogram_data <- metadata_hist_bgcs %>%
  group_by(gcf_id) %>%
  summarise(phylum_count = n_distinct(phylum))

bgcs_hist_phyla=ggplot(histogram_data, aes(x=phylum_count))+
  geom_histogram(fill="#ED496F", binwidth = 1)+
  xlab("Phyla per GCF")+
  ylab("# of GCFs")+
  theme_classic()+
  theme(axis.text = element_text(size=13.5),
        axis.title = element_text(size=15))+
  xlim(c(0,15))

#making a histogram for BGCs per GCF

proG3_gcfs=proG3_igua %>% 
  group_by(gcf_id) %>% 
  summarise(n=n())


bgcs_hist_GCs=ggplot(proG3_gcfs, aes(x=n))+
  geom_histogram(fill="#15889C", binwidth = 1)+
  xlab("BGCs per GCF")+
  ylab("# of GCFs")+
  theme_classic()+
  theme(axis.text = element_text(size=13.5),
        axis.title = element_text(size=15))+
  xlim(c(0,15))+
  scale_y_continuous(breaks = seq(0,30000, 5000))

#making the size reduction area plot


Step=c(1:4)
n=c(100, (285425/2829071)*100, (180920/2829071)*100, (56960/2829071)*100)

df_bgcs=data.frame(Step, n)

df_interp <- data.frame(
  Step = seq(1, 4, by = 0.0001)  # finer steps
) %>%
  mutate(n = approx(df_bgcs$Step, df_bgcs$n, xout = Step)$y)

df_interp$StepGroup <- cut(
  df_interp$Step,
  breaks = c(0, 1, 2, 3, 4),
  labels = c("BGCs", "FRs", "NRs", "GCFRs"),
  right = TRUE
)

# Define colors for each StepGroup
MyColours <- precols[1:3]

# Plot
bgcs_area_size=ggplot() +
  geom_area(data = df_interp, aes(x = Step, y = n, fill = StepGroup),alpha=0.45,
            show.legend = FALSE) +
  geom_line(data = df_bgcs, aes(x = Step, y = n), colour = "grey60", size = 2) +
  geom_point(data = df_bgcs, aes(x = Step, y = n), colour = "grey60", size = 3) +
  geom_point(data = df_bgcs, aes(x = Step, y = n), colour = "white", size = 1.5) +
  scale_fill_manual(values = MyColours, name = "Step Intervals") +
  scale_x_continuous(breaks = 1:4, labels = c("BGCs", "FRs", "NRs", "GCFRs")) +
  ylab("% of initial dataset size") +
  theme_classic() +
  theme(
    axis.title.x = element_blank(),
    axis.text = element_text(size = 13.5),
    axis.title = element_text(size = 15),
    legend.text = element_text(size = 13.5),
    legend.title = element_text(size = 15)
  )


#arranging all plots in a panel

fig_3=ggarrange(
  ggarrange(bgcs_hist_GCs, pro_hist_GCs, sec_hist_GCs, align="hv",
            ncol = 3, labels = c("A1", "B1", "C1"), label.y = 1, vjust = 1.3, hjust = -0.1),
  ggarrange(bgcs_area_size, pro_area_size, sec_area_size, align="hv",
            ncol = 3, labels = c("A2", "B2", "C2"), label.y = 1.08, vjust = 1, hjust = -0.1),
  ggarrange(bgcs_umap,pro_umap, sec_umap,
            align = "hv", ncol = 3,
            labels = c("A3", "B3", "C3"),label.y = 1, vjust = 1.3, hjust = -0.1),
  ggarrange(bgcs_hist_phyla, pro_hist_phyla, sec_hist_phyla, 
            align = "hv", ncol = 3,
            labels = c("A4", "B4","C4"), label.y = 1, vjust = 1.3, hjust = -0.1),
  align = "hv", heights = c(0.5,0.5,2,0.5), nrow = 4
)
ggsave("figure3.png", plot = fig_3, device = "png", dpi = 350, height = 17, width = 23)

ggsave("figure3.svg", plot = fig_3, device = "svg", dpi = 350, height = 17, width = 23)

```


```{r, results='hide', eval=FALSE, echo=TRUE}
###SUPPLEMENTARY FIGURE 1 MASTER CODE

#Setting the colour palette

bright <- color("bright")

#Loading AMI data

AMI_benchmark_data <- read.csv("AMI_benchmark_data.tsv", sep="")

#Creating AMI plot

p1=ggplot(AMI_benchmark_data,aes(x=distance,y=ARIs,group=group, color=group))+
  geom_line(lwd=1)+ylab("AMI")+xlab("Clustering distance")+
  theme_classic()+
  scale_color_manual(values=c(bright(7)))+  
  labs(color="Linkage method")+
  new_scale_color()+
  geom_hline(aes(yintercept=0.8389638,
                 color="glocal"), linetype=2)+
  geom_hline(aes(yintercept=0.8364521,
                 color="global"), linetype=3)+
  scale_color_manual(values=c("black", "black"))+
  labs(color="BiG-SCAPE")+
  theme(axis.title = element_text(size=15),
        axis.text = element_text(size=13.5),
        legend.text = element_text(size=13.5),
        legend.title = element_text(size=15))

#Loading NMI data

NMI_benchmark_data <- read.csv("NMI_benchmark_data.tsv", sep="")

#Creating NMI plot

p2=ggplot(NMI_benchmark_data,aes(x=distance,y=ARIs,group=group, color=group))+
  geom_line(lwd=1)+ylab("NMI")+xlab("Clustering distance")+
  theme_classic()+
  scale_color_manual(values=c(bright(7)))+  
  labs(color="Linkage method")+
  new_scale_color()+
  geom_hline(aes(yintercept=0.936672,
                 color="glocal"), linetype=2)+
  geom_hline(aes(yintercept=0.9347541,
                 color="global"), linetype=3)+
  scale_color_manual(values=c("black", "black"))+
  labs(color="BiG-SCAPE")+
  theme(axis.title = element_text(size=15),
        axis.text = element_text(size=13.5),
        legend.text = element_text(size=13.5),
        legend.title = element_text(size=15))

#Loading purity data

purity_benchmark_data <- read.csv("purity_benchmark_data.tsv", sep="")

#Creating purity plot

p3=ggplot(purity_benchmark_data,aes(x=distance,y=purity,group=group, color=group))+
  geom_line(lwd=1)+ylab("Purity")+xlab("Clustering distance")+
  theme_classic()+
  scale_color_manual(values=c(bright(7)))+  
  labs(color="Linkage method")+
  new_scale_color()+
  geom_hline(aes(yintercept=0.7809798,
                 color="glocal"), linetype=2)+
  geom_hline(aes(yintercept=0.7707736,
                 color="global"), linetype=3)+
  scale_color_manual(values=c("black", "black"))+
  labs(color="BiG-SCAPE")+
  theme(axis.title = element_text(size=15),
        axis.text = element_text(size=13.5),
        legend.text = element_text(size=13.5),
        legend.title = element_text(size=15))

#Loading GCF data

GCFs_benchmark_data <- read.csv("GCFs_benchmark_data.tsv", sep="")

#Creating GCF plot

p4=ggplot(GCFs_benchmark_data,aes(x=distance,y=GCFs,group=group, color=group))+
  geom_line(lwd=1)+ylab("Number of GCFs")+xlab("Clustering distance")+
  theme_classic()+
  scale_color_manual(values=c(bright(7)))+  
  labs(color="Linkage method")+
  new_scale_color()+
  geom_hline(aes(yintercept=81,
                 color="glocal"), linetype=2)+
  geom_hline(aes(yintercept=80,
                 color="global"), linetype=3)+
  scale_color_manual(values=c("black", "black"))+
  labs(color="BiG-SCAPE")+
  theme(axis.title = element_text(size=15),
        axis.text = element_text(size=13.5),
        legend.text = element_text(size=13.5),
        legend.title = element_text(size=15))

#Creating the panel figure

ggarrange(p1,p2,p3,p4,
          nrow = 2, ncol=2, align = "hv",
          labels = c("A", "B", "C", "D"))
```

```{r, results='hide', eval=FALSE, echo=TRUE}
###SUPPLEMENTARY FIGURE 2 MASTER CODE

#Filtering the original BiG-SCAPE manually curated dataset to generate the final manually curated dataset to be used

curated=read_excel("41589_2019_400_MOESM31_ESM.xlsx")
curated=curated[-c(288,289,290,291,292,293,294,308,309,310),]
sep_curated=curated %>%
  separate_rows(`MIBiG accession`, sep = ",")
sep_curated$`MIBiG accession`=str_trim(sep_curated$`MIBiG accession`, side=c("both"))
sep_curated=data.frame(sep_curated$`MIBiG accession`,sep_curated$Group)
colnames(sep_curated)=c("mibig","group")
ann_curated=data.frame(curated$Class, curated$Group)
colnames(ann_curated)=c("class","group")
BGCs=c("BGC0000015", "BGC0000095","BGC0000322",
       "BGC0000462","BGC0000629","BGC0000875",
       "BGC0000945", "BGC0001081", "BGC0001147",
       "BGC0001208", "BGC0001232", "BGC0001347")
b=length(BGCs)
temp1=0
for (y in 1:b){
  temp1=c(temp1,c(which(sep_curated$mibig==BGCs[y])))
}
sep_curated=sep_curated[-temp1,]

#Importing the BiG-SCAPE clustering of the manually curated dataset and filtering out records removed from the original curated dataset

big_scape_glocal <- read.delim("supplementary_table_S2.tsv", header=FALSE, comment.char="#", stringsAsFactors=TRUE)
colnames(big_scape_glocal)=c("mibig","GCF")
big_scape_glocal$mibig=sapply(big_scape_glocal$mibig, function(x)
  sub(".1", "",x,fixed=TRUE))

glocal=big_scape_glocal %>%
  group_by(GCF) %>%
  filter(n() >=2)

BGCs=c("BGC0000015", "BGC0000095","BGC0000322",
       "BGC0000462","BGC0000629","BGC0000875",
       "BGC0000945", "BGC0001081", "BGC0001147",
       "BGC0001208", "BGC0001232", "BGC0001347")
n=length(BGCs)
temp=0

for (i in 1:n){
  temp=c(temp,c(which(glocal$mibig==BGCs[i])))
}
glocal_f=glocal[-temp,]

merged_data2=merge(sep_curated,glocal_f,by="mibig", all.x = TRUE)

data3=merged_data2 %>%
  group_by(group, GCF) %>% 
  summarise(n=n()) 

data3$row=row.names(data3)

#Creating annotations

ann_merge=data3 %>% inner_join(ann_curated,
                                     by="group")

ann_merge=ann_merge[!duplicated(ann_merge$row),]
ann_row=data.frame(ann_merge$group,ann_merge$class)
ann_row=ann_row[!duplicated(ann_row$ann_merge.group),]
row.names(ann_row)=ann_row$ann_merge.group
ann_row2=data.frame(ann_row$ann_merge.class)
row.names(ann_row2)=row.names(ann_row)
colnames(ann_row2)=c("Class")

#Sorting the columns

contingency.table <- table(merged_data2$GCF, merged_data2$group)
my.truth <- names(sort(table(merged_data2$group), decreasing = T))
gecco.clusters <- names(sort(table(merged_data2$GCF), decreasing = T))

yeet <- function(mymat, ref){
  find.clusters <- c()
  used.clusters <-c()
  # loop through reference list
  for (i in 1:length(ref)){
    clus <- ref[i]
    # take the column of the contintingency table that matches the current reference GCF
    # finds the row that yields the GCF with maximum overlap
    # if there are ties, R takes the first one
    clus.bgcs <- rownames(mymat)[which.max(mymat[,which(colnames(mymat)==clus)])]
    # if that particular maximum GCF hasn't been included yet...
    # (this is done to deal with ties)
    if (!(clus.bgcs%in%used.clusters)){
      # append the maximum GCF to our final list of clusters, as well as out "used" GCF list
      find.clusters <- c(find.clusters, clus.bgcs)
      used.clusters <- c(used.clusters, clus.bgcs)
    }
  }
  for (i in 1:length(rownames(mymat))){
  clus.missing <- rownames(mymat)[i]
  # if the GCF isn't included in our final list...
  if (!(clus.missing%in%used.clusters)){
    # add it
    used.clusters <- c(used.clusters, clus.missing)
    find.clusters <- c(find.clusters, clus.missing)
  }
}
# return ordered GCFs
return(find.clusters)
}


order.big <- yeet(mymat = contingency.table, ref = my.truth)
final <- contingency.table
final <- final[match(order.big, rownames(final)),]
order.gecco <- yeet(mymat = t(final), ref = rownames(final))
final <- final[,match(order.gecco, colnames(final))]

#Creating the final heatmap

p8=as.ggplot(pheatmap(t(final), color = hcl.colors(50, "YlGnBu"), border_color = F,cluster_rows = F,
         cluster_cols = F,
         show_colnames = FALSE,fontsize = 5, annotation_row = ann_row2,
         annotation_colors = ann_colors, annotation_names_row = FALSE))
p8

```

```{r, results='hide', eval=FALSE, echo=TRUE}
###SUPPLEMENTARY FIGURE 3 MASTER CODE

##Panel A

#Loading BiG-SCAPE clustering of the Streptomyces dataset

big_scape <- read.delim("supplementary_table_S5.tsv", header=FALSE, comment.char="#")
colnames(big_scape)=c("BGC","GCF_b")

#Removing singleton GCFs

big_scape=big_scape%>%
  group_by(GCF_b) %>%
  filter(n() >=2)

#Loading the IGUA clustering of the Streptomyces dataset

igua_strepto <- read.delim("supplementary_table_S6.tsv")
igua=data.frame(igua_strepto$cluster_id, igua_strepto$gcf_id)
colnames(igua)=c("BGC","GCF_h")

#Removing singleton GCFs

igua=igua%>%
  group_by(GCF_h) %>%
  filter(n() >=2)

#Merging the two datasets

merged_data=merge(big_scape,igua,by="BGC", all.y = TRUE)

data2=merged_data %>% 
  group_by(GCF_h, GCF_b) %>% 
  summarise(n=n())

#Creating the alluvial plot

p10=ggplot(data2,
       aes(y = n, axis1 = GCF_h, axis2 = GCF_b)) +
  geom_alluvium(width = 1/12, aes(fill="dodgerblue"), show.legend = FALSE) +
  scale_x_discrete(limits = c("GCF_h", "GCF_b"), expand = c(.05, .05)) +
  scale_fill_manual(values=c("dodgerblue"))+
  theme_classic()+
  theme(axis.line.y = element_blank(), axis.text.y = element_blank(),
        axis.ticks.y = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_text(size = 10))+
  scale_x_continuous(breaks = 1:2, labels=c("IGUA","BiG-SCAPE"))

p10

##Panel B

#Importing IGUA clustering of the Streptomyces dataset

igua_strepto <- read.delim("supplementary_table_S6.tsv")

#Filtering singleton GCFs

data=igua_strepto %>%
  group_by(gcf_id) %>%
  summarise(n=n())
#Importing the BiG-SCAPE clustering of the Streptomyces dataset

big_scape <- read.delim("supplementary_table_S5.tsv", header=FALSE, comment.char="#")
colnames(big_scape)=c("BGC","GCF_b")

#Filtering singleton GCFs

data2=big_scape %>%
  group_by(GCF_b) %>%
  summarise(n=n())

#Creating the histogram

from <- c(xmin = 0, xmax = 15, ymin = 0, ymax = 3500)
to <- c(xmin = 200, xmax = 480, ymin = 1000, ymax = 3600)

p12=ggplot()+
  geom_histogram(data=data2, aes(x=n, fill="b", color="b"),
                 binwidth = 1,show.legend=TRUE, size=0.3, alpha=0.9)+
  geom_histogram(data=data, aes(x=n, fill="p", color="p"),
                 binwidth = 1, show.legend = TRUE, size=0.3, alpha=0.9)+
  xlab("BGCs per GCF")+
  ylab("Number of GCFs")+
  theme_classic()+
  scale_color_manual(name="Clustering tool", values=c("b"="black", "p"="black"), labels=c("b"="BiG-SCAPE", "p"="IGUA"))+
  scale_fill_manual(name="Clustering tool", values=c("b"="#33BBEE", "p"="#009988"), labels=c("b"="BiG-SCAPE", "p"="IGUA"))+
  geom_magnify(data=data, from=from, to=to, shadow = FALSE,
               axes = "xy",
               corners = 0.05,
               proj = "facing",
               linewidth = 0.35,
               linetype = 1,
               proj.linetype = 3)
p12

##Panel C

#Importing the BiG-SCAPE clustering of the Streptomyces dataset

big_scape <- read.delim("supplementary_table_S5.tsv", header=FALSE, comment.char="#")
colnames(big_scape)=c("BGC","GCF_b")

#Removing singleton GCFs
big_scape=big_scape%>%
  group_by(GCF_b) %>%
  filter(n() >=2)

#Importing the IGUA clustering of the Streptomyces dataset

igua_strepto <- read.delim("supplementary_table_S6.tsv")
igua=data.frame(igua_strepto$cluster_id, igua_strepto$gcf_id)
colnames(igua)=c("BGC","GCF_h")

#Filtering singleton GCFs

igua=igua%>%
  group_by(GCF_h) %>%
  filter(n() >=2)

#Merging the two datasets

merged_data=merge(big_scape,igua,by="BGC", all.x = TRUE)

data2=merged_data %>%
  group_by(GCF_b, GCF_h) %>% 
  summarise(n=n()) 

data2=data2[order(data2$n, decreasing = TRUE),]

overlap=data2$GCF_b[1:100]
df=data.frame(overlap)
colnames(df)=c("GCF_b")

merged=merge(df,merged_data, by="GCF_b", all.x = TRUE)
cont=table(merged$GCF_b,merged$GCF_h)

#Creating the heatmap


p9=as.ggplot(pheatmap(cont, color = hcl.colors(50, "YlGnBu"), border_color = F,cluster_rows = TRUE,
                      cluster_cols = TRUE, clustering_distance_rows = "euclidean",
                      show_colnames = FALSE,fontsize = 8, annotation_names_row = FALSE,
                      show_rownames = FALSE,
                      heatmap_legend_param = list(title = "# of BGCs")))

p9

#NOTE: clustering dendrograms were removed using Inkscape

```

```{r, results='hide', eval=FALSE, echo=TRUE}
###SUPPLEMENTARY FIGURE 4 MASTER CODE

#Creating the dataframe - note: this is the equivalent of supplementary table 8

sample_size=c(rep(1000, 3), rep(2500, 3), rep(5000,3), rep(10000,3))
time=c(27.8, 28, 35.6,
       71, 73, 76,
       190, 187, 182,
       497, 477, 470)

scatter_df=data.frame(sample_size,time)

#Creating the plot

p11=ggplot(scatter_df, aes(x=sample_size, y=time))+
  geom_point(color="#4477AA", alpha=0.6, aes(size=1.5), show.legend = FALSE)+
  theme_classic()+
  xlab("Number of BGCs")+
  ylab("Real time (sec)")+
  geom_smooth(method = "glm", colour="black", lwd=0.5,
              linetype=2, se=FALSE)+
  theme(axis.title = element_text(size=15),
        axis.text = element_text(size = 13.5))
p11
```

```{r, results='hide', eval=FALSE, echo=TRUE}
###SUPPLEMENTARY FIGURE 5 MASTER CODE

##Panel A

#Importing the Seurat object associated with the proG3 dataset

pfams <- readRDS("pfams_proG3.RDS")

#Making the clustree figure

clustree(pfams@meta.data, prefix = "peaks_snn_res.") 

##Panel B

#Importing the BGC labels

labels <- read.delim("proG3_labels.tsv", header=FALSE)
colnames(labels)=c("BGC", "GCF")

#Importing the biosynthetic class annotation associated with the proG3 dataset

classes <- read.delim("biosynthetic_classes.tsv", header=FALSE)
colnames(classes)=c("BGC","Biosynthetic_class")

#Sorting classes

classes$Biosynthetic_class=sapply(classes$Biosynthetic_class, function(x)
  sub("saccharide", "Saccharide",x,fixed=TRUE))
classes$Biosynthetic_class=sapply(classes$Biosynthetic_class, function(x)
  sub("terpene", "Terpene",x,fixed=TRUE))
classes$Biosynthetic_class=sapply(classes$Biosynthetic_class, function(x)
  sub("NRPS", "NRP",x,fixed=TRUE))
classes$Biosynthetic_class=sapply(classes$Biosynthetic_class, function(x)
  sub("NRP;Polyketide", "NRP-PKS",x,fixed=TRUE))
classes$Biosynthetic_class=sapply(classes$Biosynthetic_class, function(x)
  sub("Polyketide", "PKS",x,fixed=TRUE))
classes$Biosynthetic_class=sapply(classes$Biosynthetic_class, function(x)
  sub("PKS;Saccharide", "Others",x,fixed=TRUE))
classes$Biosynthetic_class=sapply(classes$Biosynthetic_class, function(x)
  sub("RiPP;Terpene", "Others",x,fixed=TRUE))
classes$Biosynthetic_class=sapply(classes$Biosynthetic_class, function(x)
  sub("Unknown", "Others",x,fixed=TRUE))
classes$Biosynthetic_class=sapply(classes$Biosynthetic_class, function(x)
  sub("mixed", "Others",x,fixed=TRUE))
classes$Biosynthetic_class=sapply(classes$Biosynthetic_class, function(x)
  sub("other", "Others",x,fixed=TRUE))
classes$Biosynthetic_class=sapply(classes$Biosynthetic_class, function(x)
  sub("NRP;Saccharide", "Others",x,fixed=TRUE))
classes$Biosynthetic_class=sapply(classes$Biosynthetic_class, function(x)
  sub("PKS;Terpene", "Others",x,fixed=TRUE))
classes$Biosynthetic_class=sapply(classes$Biosynthetic_class, function(x)
  sub("NRP;Saccharide", "Others",x,fixed=TRUE))
classes$Biosynthetic_class=sapply(classes$Biosynthetic_class, function(x)
  sub("PKS;Terpene", "Others",x,fixed=TRUE))
classes$Biosynthetic_class=sapply(classes$Biosynthetic_class, function(x)
  sub("Alkaloid;Terpene", "Others",x,fixed=TRUE))
classes$Biosynthetic_class=sapply(classes$Biosynthetic_class, function(x)
  sub("Alkaloid;NRP-PKS", "Others",x,fixed=TRUE))
classes$Biosynthetic_class=sapply(classes$Biosynthetic_class, function(x)
  sub("Saccharide;Terpene", "Others",x,fixed=TRUE))

#Importing the IGUA clustering of the proG3 dataset

proG3_igua=read.delim('supplementary_table_S11.tsv', header=TRUE)
colnames(proG3_igua)[1]="BGC"

#Attaching the class annotation to the IGUA clustering dataframe

data=merge(proG3_igua, classes, by="BGC")

#Creating a Majority count column

df=data %>% 
  add_count(gcf_id, Biosynthetic_class) %>% 
  group_by(gcf_id) %>% 
  mutate(Majority=Biosynthetic_class[n==max(n)][1]) %>% 
  select(-n)

#Merging to the labels

data2=merge(labels, df, by="BGC")

data3=data2[!duplicated(data2),]

rownames(data3)=data3$BGC

#Attaching biosynthetic class annotation with majority count column as metadata to the Seurat object

meta_data=pfams@meta.data


pfams=AddMetaData(pfams, metadata = data3)

Idents(object = pfams) <- "Majority"

#Creating the UMAP plot


p17_5=DimPlot(pfams, label=FALSE, cols=c('Alkaloid' = "#EF6C00", 'NRP' = "#884EA0", 'Others' = "grey", 'PKS' = "#1E88E5", 'NRP-PKS' = "palegreen",
            'RiPP' = "#FDD835", 'Saccharide' = "#EC407A", 'Terpene' = "#009688"),
            alpha=0.5)

p17_5

```

```{r, results='hide', eval=FALSE, echo=TRUE}
###SUPPLEMENTARY FIGURE 6 MASTER CODE

##Panel A

#Importing the secretion systems Seurat object

pfams_sec <- readRDS("pfams_secretion.RDS")

#Creating clusterings for clustree

res = seq.int(0.1, 0.9, 0.05) 
for (i in res){
  pfams_sec = FindClusters(object = pfams_sec, resolution = i, algorithm = 3)
}

#Plottting the clustree figure

clustree(pfams_sec@meta.data, prefix = "peaks_snn_res.") 

##Panel B

#Importing the secretion systems Seurat object

pfams_sec <- readRDS("pfams_secretion.RDS")

#Setting the colour palette

precols = c("#15889C", "#ED496F", "#8E1382", "#FFB300", "#FF6F00", "#43A047",
            "#2D62A3", "#1F7530", "#573794", "#28AAE1", "#870F0F", "#24b177",
            "#DDDDDD", "#7DCEA0", "#85C1E9", "#EB984E", "#EAD9D5", "#03CD4A",
            "#CDDC39", "#E0594B", "#C76CDE", "#8D6E63", "#486FF7", "#6300B5",
            "#88E200", "#012824", "#0D3290", "#A347FB", "#54FC7A", "#EB1388",
            "#B0978D", "#FE52CF", "#83F1F6", "#F1F847", "#2B1DFC", "#6C6F15", 
            "#6CA05C", "#7788CD", "#F502F3", "#0DC290", "#FA0E03", "#3CAA0A",
            "#BEFC8D", "#08F8EB", "#B1CD3F", "#D6A5FA", "#CE606C", "#AB1EBA",
            "#6ECC9F", "#054DDC", "#854F49", "#F22B21", "#3A0E43", "#225805",
            "#37D160", "#E4B974", "#A8BADE", "#47EDD1", "#F47A92", "#9106EB",
            "#81AA20", "#D7FDFD", "#5DEB2E", "#F82745", "#6435E0", "#027FFE",
            "#8E3101", "#16F648", "#1C15BC", "#8BE46E", "#8D6FA0", "#E68FC6",
            "#058CA9", "#9E018A", "#BDFD0B", "#B22760", "#2BF49F", "#CB9348",
            "#9D8303", "#C251A1", "#46ADAF", "#A3E3AF", "#22BB34", "#6EA3FA",
            "#260374", "#1C3854", "#405D37", "#C21DF3", "#FCEA92", "#537F88",
            "#FD4C18", "#F2D71E")

#Importing and merging the files

secretion_metadata <- read.delim("single_locus_clusters_renamed.tsv", header=FALSE)
sec_igua=read.delim('supplementary_table_S13.tsv', header=TRUE)
labels_2 <- read.delim("sec_labels.tsv", header=FALSE)

colnames(sec_igua)[1]="secretion"
colnames(labels_2)=c("secretion","gcf_id")
colnames(secretion_metadata)=c("contig_id","secretion","start","end","major_secretion_system_type",
                                    "minor_secretion_system_type","NCBI_nucleotide_accession")

metadata_igua=merge(secretion_metadata, sec_igua, by="secretion")

#Creating a majority-class column

metadata_igua=metadata_igua %>% 
  add_count(gcf_id, major_secretion_system_type) %>% 
  group_by(gcf_id) %>% 
  mutate(Majority=major_secretion_system_type[n==max(n)][1]) %>% 
  select(-n)

metadata_secretion=merge(labels_2, metadata_igua, by="secretion", all.x=TRUE)

row.names(metadata_secretion)=metadata_secretion$secretion

#Attaching dataframe as metadata

pfams_sec=AddMetaData(pfams_sec, metadata = metadata_secretion)

#Creating the plot

DimPlot(object = pfams_sec, label = FALSE, cols=precols, alpha=0.5,
        group.by = "Majority", pt.size = 1)

```

```{r, results='hide', eval=FALSE, echo=TRUE}
###SUPPLEMENTARY FIGURE 7 MASTER CODE

#Importing the secretion systems Seurat object

pfams_sec <- readRDS("pfams_secretion.RDS")

#Setting the colour palette

precols = c("#15889C", "#ED496F", "#8E1382", "#FFB300", "#FF6F00", "#43A047",
            "#2D62A3", "#1F7530", "#573794", "#28AAE1", "#870F0F", "#24b177",
            "#DDDDDD", "#7DCEA0", "#85C1E9", "#EB984E", "#EAD9D5", "#03CD4A",
            "#CDDC39", "#E0594B", "#C76CDE", "#8D6E63", "#486FF7", "#6300B5",
            "#88E200", "#012824", "#0D3290", "#A347FB", "#54FC7A", "#EB1388",
            "#B0978D", "#FE52CF", "#83F1F6", "#F1F847", "#2B1DFC", "#6C6F15", 
            "#6CA05C", "#7788CD", "#F502F3", "#0DC290", "#FA0E03", "#3CAA0A",
            "#BEFC8D", "#08F8EB", "#B1CD3F", "#D6A5FA", "#CE606C", "#AB1EBA",
            "#6ECC9F", "#054DDC", "#854F49", "#F22B21", "#3A0E43", "#225805",
            "#37D160", "#E4B974", "#A8BADE", "#47EDD1", "#F47A92", "#9106EB",
            "#81AA20", "#D7FDFD", "#5DEB2E", "#F82745", "#6435E0", "#027FFE",
            "#8E3101", "#16F648", "#1C15BC", "#8BE46E", "#8D6FA0", "#E68FC6",
            "#058CA9", "#9E018A", "#BDFD0B", "#B22760", "#2BF49F", "#CB9348",
            "#9D8303", "#C251A1", "#46ADAF", "#A3E3AF", "#22BB34", "#6EA3FA",
            "#260374", "#1C3854", "#405D37", "#C21DF3", "#FCEA92", "#537F88",
            "#FD4C18", "#F2D71E")

#Running with default clustering threshold

pfams_sec <- FindClusters(object = pfams_sec, verbose = FALSE, algorithm = 3)

#Creating the UMAP plot

DimPlot(object = pfams_sec, label = TRUE, cols=precols, alpha=0.5,
        pt.size = 1, label.size = 6) + NoLegend()

```

```{r, results='hide', eval=FALSE, echo=TRUE}
###SUPPLEMENTARY FIGURE 8 MASTER CODE

##Panel A

#Importing prophage Seurat object

prophage=readRDS('pfams_prophage.RDS')

#Running clustering at different thresholds

res = seq.int(0.1, 0.9, 0.05) 
for (i in res){
  prophage = FindClusters(object = prophage, resolution = i, algorithm = 3)
}

#Making the clustree figure

clustree(prophage@meta.data, prefix = "peaks_snn_res.") 

##Panel B

#Importing the prophage Seurat object

pfams_pro = readRDS("pfams_prophage.RDS")

#Setting the colour scheme

precols = c("#15889C", "#ED496F", "#8E1382", "#FFB300", "#FF6F00", "#43A047",
            "#2D62A3", "#1F7530", "#573794", "#28AAE1", "#870F0F", "#24b177",
            "#DDDDDD", "#7DCEA0", "#85C1E9", "#EB984E", "#EAD9D5", "#03CD4A",
            "#CDDC39", "#E0594B", "#C76CDE", "#8D6E63", "#486FF7", "#6300B5",
            "#88E200", "#012824", "#0D3290", "#A347FB", "#54FC7A", "#EB1388",
            "#B0978D", "#FE52CF", "#83F1F6", "#F1F847", "#2B1DFC", "#6C6F15", 
            "#6CA05C", "#7788CD", "#F502F3", "#0DC290", "#FA0E03", "#3CAA0A",
            "#BEFC8D", "#08F8EB", "#B1CD3F", "#D6A5FA", "#CE606C", "#AB1EBA",
            "#6ECC9F", "#054DDC", "#854F49", "#F22B21", "#3A0E43", "#225805",
            "#37D160", "#E4B974", "#A8BADE", "#47EDD1", "#F47A92", "#9106EB",
            "#81AA20", "#D7FDFD", "#5DEB2E", "#F82745", "#6435E0", "#027FFE",
            "#8E3101", "#16F648", "#1C15BC", "#8BE46E", "#8D6FA0", "#E68FC6",
            "#058CA9", "#9E018A", "#BDFD0B", "#B22760", "#2BF49F", "#CB9348",
            "#9D8303", "#C251A1", "#46ADAF", "#A3E3AF", "#22BB34", "#6EA3FA",
            "#260374", "#1C3854", "#405D37", "#C21DF3", "#FCEA92", "#537F88",
            "#FD4C18", "#F2D71E")

#Importing and sorting the metadata

prophage_metadata <- read.delim("prophage_metadata.tsv", header=TRUE)
labels <- read.delim("prophage_labels.tsv", header=FALSE)
colnames(labels)=c("prophage","GCF")
colnames(prophage_metadata)[1]="prophage"
metadata_prophage=merge(labels, prophage_metadata, by="prophage", all.x=TRUE)
rownames(metadata_prophage)=metadata_prophage$prophage

#Keeping the top 10 phyla

top10 = metadata_prophage %>%
  group_by(phylum) %>% 
  summarise(n=n()) %>% 
  top_n(11)

#Modifying the metadata to keep only the top 10 phyla

data3 = metadata_prophage %>% 
  filter(phylum %in% top10$phylum)

#Attaching the new metadata to the Seurat object

pfams_3=AddMetaData(pfams_pro, metadata = data3)

#Setting identity to phylum

Idents(object = pfams_3) <- "phylum"


#Creating the UMAP plot

DimPlot(pfams_3, na.value = "white", raster=FALSE, cols = precols,
        alpha = 0.5)


```

